### 2022-01-05 manuscript/coverLetter/Letter.tex
updated and prepped for release and zenodo-ing

\documentclass[12pt]{article}
\usepackage[letterpaper,margin=1cm,bottom=2cm,top=3cm,nohead]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\newcommand{\email}[1]{\href{mailto:#1}{\tt \textcolor{cornflower}{#1}}}
\usepackage{xcolor}
\definecolor{plum}{rgb}{0.36078, 0.20784, 0.4}
\definecolor{chameleon}{rgb}{0.30588, 0.60392, 0.023529}
\definecolor{cornflower}{rgb}{0.12549, 0.29020, 0.52941}
\definecolor{scarlet}{rgb}{0.8, 0, 0}
\definecolor{brick}{rgb}{0.64314, 0, 0}
\usepackage[%
    colorlinks=true,  %display links in different colors
    linkcolor=cornflower, %internal links
    citecolor=scarlet, %references to a bibliography
    urlcolor=chameleon %URLs
]{hyperref}
\usepackage{graphicx}
\setlength{\parskip}{0.75em}
\setlength{\parindent}{1.75em}
\usepackage{afterpage}
\usepackage[strict]{changepage}
\pagestyle{empty}

\begin{document}

\begin{minipage}[c]{8in}\vskip-2cm
\begin{flushleft}
	\begin{minipage}[c]{6cm}
		\begin{flushleft}
			\includegraphics*[width=6cm]{usc.jpg}%
		\end{flushleft}
	\end{minipage}
	\,\,
	\begin{minipage}[c]{0.25cm}
		\vrule depth 0.1\textheight
	\end{minipage}
	\,
	\begin{minipage}[c]{8cm}
		\footnotesize
		{\sffamily
			Tad Dallas \\
			Department of Biological Sciences \\
			University of South Carolina\\
			Columbia, SC 29208 \\
			\email{tdallas@mailbox.sc.edu}
		}
	\end{minipage}
\end{flushleft}
\end{minipage}

\begin{adjustwidth}{4em}{4em}
~\vskip.4cm
\hfill \today
\vskip.5cm

{\noindent Dear \textit{Ecology Letters} editorial board,} \vskip4pt

Please find the enclosed manuscript ``\texttt{insectDisease}: programmatic access to the \textit{Ecological Database of the World's Insect Pathogens}'', which we re-submit for consideration as a research article. \\

Curated databases of species interactions are instrumental to exploring and understanding the spatial distribution of species and their biotic interactions. In the process of conducting such projects, data development and curation efforts may give rise to a data product with utility beyond the scope of the original work, but which becomes inaccessible over time. We present one such example, the "Ecological Database of World's Insect Pathogens". The R data package and manuscript text have benefitted from the comments of two reviewers and editor Dr. Michael Borregaard. These comments centered on the package structure and the presentation of the data. Specifically, the data required processing, but we were unaware of the current standard practice of having a \texttt{data-raw} folder and a processing script. We have incorporated this, and made more data checks -- including the incorporation of unit tests in the pacakge -- to handle odd date issues, remove bad unicode characters, and provide some signal of potential error as these data will be augmented in the future with more records. The comments from all reviewers seemed to indicate that substantial revision was necessary, and we believe we have made substantial modifications to both manuscript text and R package structure. We thank the reviewers and the editor for their critiques. \\

We believe this article will be of interest to a broad range of ecologists interested in pest management, disease, species interactions, and more. In advance, we appreciate your consideration and assistance. Please don't hesitate to contact us with any questions.

\changepage{}{}{}{}{}{1cm}{}{}{}

\vskip1cm
\hfill
\noindent\begin{minipage}[b]{7cm}
    Sincerely,\\ ~
    Tad Dallas
\end{minipage}

\end{adjustwidth}
\end{document}